---
# Standalone working draft. This file is excluded from the book render (see
# `project: render:` in ../_quarto.yml); the book calls scale_ladder_plot()
# directly from index.qmd. Render it on its own with:
#   quarto render scale-ladder/figure-scale-ladder.qmd
# Its bibliography is held out of the book's references pending review.
title: "The AI race is a hiring race"
bibliography: figure_references_for_review.bib
format:
  pdf:
    fig-width: 7.5
    fig-height: 2.4
    keep-tex: true
---

```{r}
#| label: setup
#| include: false
source("R/scale-ladder.R")   # defines has_cairo() and scale_ladder_plot()
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      dev = if (has_cairo()) "cairo_pdf" else "pdf")
```

```{r}
#| label: fig-scale-ladder
#| fig-cap: "Author counts on flagship model reports, 2018--2025. One square is one credited person; each block's area is its author list."
scale_ladder_plot(fig_width = 7.5)
```

Eight names wrote the paper that started this [@vaswani2017attention]. Eight
years later a single flagship report credits 3,291 [@gemini2025gemini25] --- more
than the 2,932 on the ATLAS paper that announced the Higgs [@atlas2012higgs], and
past the 1,004 on LIGO's first detection [@ligo2016observation].

The series is not monotonic. Gemini 1.5 [@gemini2024gemini15] credits roughly
two thirds as many people as Gemini 1.0 [@gemini2023gemini], and DeepSeek-V3
reached the frontier on under 200 [@deepseek2024v3].

Nor is the disclosure itself stable, and the instability runs in both
directions. Where a list is published, it keeps growing after release: Gemini
1.0's contributor list stood at 939 names in December 2023 and at 1,350 by the
2025 revision, so the same report answers the question differently depending on
when you ask. Where a list is not published, the arXiv byline can invent one ---
OpenAI's GPT-5 system card carries no author list anywhere in its 61 pages, yet
its arXiv record names 483 people [@openai2025gpt5]. Anthropic
[@anthropic2024claude3], xAI [@xai2024grok2] and Gemini 3 [@google2025gemini3]
publish neither, and cannot be plotted.

## Data provenance

Every count is pinned to the original release (arXiv v1) and counts unique named
individuals; an institutional byline is not counted as a person. All sixteen rows
were checked against primary sources on 2026-08-23. `data/verification-log.md`
records what was checked against what, and flags the three rows that still want a
human eye --- GPT-4, whose v1 count is derived rather than enumerated, GPT-5,
whose count exists only in arXiv metadata, and the GPT-1 date.
