```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("scale-ladder/R/scale-ladder.R")
```

# Preface {.unnumbered}

AI is radically redefining the nature of scientific inquiry — for better and for worse. 

In Fall 2026, faculty, graduate students, and postdocs across multiple departments at the University of Florida came together to understand this urgent development. In a seminar titled *Collective [Artificial] Intelligence in Science*, we documented and discussed how AI is changing each component of our research workflow. This includes project management, idea development, data analysis, writing, and dissemination. To push our understanding, we alternated between the perspectives of an AI zealot and an AI critic — identifying key resources, consensus statements, and areas of disagreement along the way. The result is a *human-curated and written* **2027 UF Guide to AI in Science**.

The seminar began with an introduction to big-team science — its rise[@Coles2022], competitive advantages[@Coles2026], and relevance to the current moment. For instance, the paper describing OpenAI’s GPT-1 had 4 authors; the paper describing GPT-3 had 31 authors; and the paper describing GPT-5 had 483 authors (@fig-scale-ladder). Similarly, a prominent AI-led cybersecurity attack on the online community, Hugging Face, involved coordinated action among thousands of AI agents[@openai2026huggingface]. After reviewing these developments, the course itself became a large-scale collaboration, dedicated to understanding how AI is changing the nature of science.

```{r}
#| label: fig-scale-ladder
#| fig-cap: "Credited co-authors on flagship AI model reports from OpenAI, Google, and Meta, 2018--2025. One square is one credited person; each block's area is its author list."
#| fig-width: 7.5
#| fig-height: 2.4
#| fig-align: left
scale_ladder_plot(fig_width = 7.5)
```


Each week followed a consistent structure (@fig-seminar-structure). First, we would select a component of the research pipeline to examine (e.g., data analysis). Second, we would spend one week compiling resources for understanding the emerging role of AI in that part of the research pipeline. Third, we would engage in peer-to-peer teaching — all with the goal of identifying the top 3 resources to include in the guide. Last, we would use a live online voting platform to solicit, debate, and vote upon key takeaways for the week ([https://pol.is/](https://pol.is/signin)). The result is the **2027 UF Guide to AI in Science**: a summary of our top resources, takeaways, and disagreements.

![Structure of the UF Fall 2026 seminar, titled Collective \[Artificial\] Intelligence in Science.](ClassStructure.png){#fig-seminar-structure fig-align="left" width="85%"}
