```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("week-summary/R/week-summary.R")
```

# Project Management: Using AI {#sec-project-management-using}

> *How is AI changing the way research teams plan, coordinate, and keep track of their work?*

**Lens:** receptive — what AI makes newly possible at this stage of the research pipeline.

```{r}
#| label: fig-week-01-summary
#| fig-cap: "What the seminar concluded about using AI for project management. Left: statements put to the group, ordered from strongest agreement to strongest disagreement. Right: the three resources voted forward, each linked by QR code and credited to the student who surfaced it."
#| fig-width: 12
#| fig-height: 5.5
#| fig-align: center
week_summary_plot(week = 1, fig_width = 12, fig_height = 5.5)
```

::: {.callout-note appearance="simple"}
**Wireframe.** @fig-week-01-summary is drawn from placeholder rows in
`week-summary/data/week-01-*.csv`. The bars, titles, and names are layout
stand-ins, not results. Replace the two CSVs and the figure fills itself in.
:::

## The Case

*[Open with one concrete situation — a real project, told in a paragraph or two.
Something a reader in this seminar would recognize: a multi-site study whose
coordination overhead outgrew the science, a lab notebook nobody kept current, a
grant timeline that slipped. Establish what the coordination problem actually
cost before naming any tool.]*

*[Then the turn into the week: what an AI-assisted version of this situation
would look like, and which of the questions below it raises. This paragraph is
the transition — it should leave the reader holding a specific question that the
resources and the votes then answer.]*

## Top Resources

The three resources the seminar voted to carry forward, in the order they appear
in @fig-week-01-summary.

### 1. *[Resource title]*

*[Author(s), year — paper, tool, thread, talk, or documentation]* · **Identified by [student name]**

**What it is:**

**Why it made the cut:**

**Who it is for:**

### 2. *[Resource title]*

*[Author(s), year]* · **Identified by [student name]**

**What it is:**

**Why it made the cut:**

**Who it is for:**

### 3. *[Resource title]*

*[Author(s), year]* · **Identified by [student name]**

**What it is:**

**Why it made the cut:**

**Who it is for:**

## Where We Agreed

The statements that reached consensus in the Polis session — the upper bars of
@fig-week-01-summary.

*[One short paragraph framing what the agreement has in common. If the
consensus statements share a shape — say, agreement on disclosure but not on
delegation — name it here rather than leaving the reader to infer it.]*

1. *[consensus statement — % agreeing]*
2. *[consensus statement — % agreeing]*

## Where We Did Not

The statements that split the group — the lower bars of @fig-week-01-summary.

*[One short paragraph on the nature of the disagreement. Whether it is empirical
(we disagree about what is true) or normative (we disagree about what is
acceptable) is the useful distinction, because only the first kind is settled by
evidence.]*

- *[position vs. counter-position — and what evidence would move it]*
- *[position vs. counter-position — and what evidence would move it]*

## Back to The Case

*[Return to the opening situation and re-run it against what the week
established. What would the group's consensus have changed about how that
project was managed? What would it not have changed? Where the group disagreed,
say plainly that the case does not settle it either.]*

*[Close on the handoff to @sec-project-management-critiquing — this chapter took
the receptive lens; the next one takes the critical one on the same stage of the
pipeline.]*

## Open Questions

- *[question the week raised but could not answer]*
