```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("week-summary/R/week-summary.R")
source("checkin-workflow/R/checkin-workflow.R")
```

# Ideation and Project Management: Using AI {#sec-project-management-using}
```{r}
#| label: fig-week-01-summary
#| fig-cap: "Overview of Week 2 of the Fall 2026 Collective [Artificial] Intelligence Seminar. Left: top discussion points, along with estimated level of agreement among seminar participants. Right: three most popular resources identified by seminar participants."
#| fig-width: 12
#| fig-height: 5.5
#| fig-align: center
week_summary_plot(week = 1, fig_width = 12, fig_height = 5.5)
```

Many researchers' first contact with AI tools is in the realm of ideation or project management. For example, one seminar participant noted that they have their research trainees submit weekly project updates to a repository shared with remote deployment of the LLM, Claude Opus (@fig-checkin-workflow, panel A). This allowed them to later prompt the model with questions about past conversations, challenges, decisions, and solutions (@fig-checkin-workflow, panel B). This same seminar participant noted that this allowed them to efficiently take stock of the trainee's progress, have a conversation with the LLM about potential future directions, and generate a simulated figure to send back to the trainee to convey next steps (@fig-checkin-workflow, panel C).

```{r}
#| label: fig-checkin-workflow
#| fig-cap: "A seminar participant shared an LLM-assisted workflow to compile past meeting notes from their trainee, review progress, and suggest future directions."
#| fig-width: 12
#| fig-height: 11.63
#| fig-align: center
checkin_workflow_plot()
```

As illustrated by the most popular weekly resources, AI increasingly being used to expand ideation and project management capabilities (@fig-week-01-summary). For example, as early as 2025, researchers documented that LLMs were capable of generating research ideas that were judged more novel (but slightly less feasible) than human experts[@ICLR2025_ea94957d]. The performance of these LLMs can be further improved via causal graphs. More specifically, LLMs can efficiently extract causal relation pairs from published articles. This allows the subsequent deployment of algorithms designed to predict missing links in this causal chain[@tong2024automating]. Similarly, LLMs can be used to *generate* theories based on pre-existing swaths of qualitative data [@schmitt2016] – an AI-for-data-analysis theme we continued to explore later in the seminar (see Chapter X).

# Areas of consensus – and dissensus

As expected, seminar discussion was marked by both areas of consensus and dissensus. In terms of consensus, there was broad agreement that LLMs can assist with ideation, project management, and the organization of qualitative sources of information. There was also broad agreement that this is changing the nature of competition in science – placing more emphasis on who can better leverage these tools for ideation and execution. Yet, there was clear disagreement about how far things should go. Seminar participants strongly disagreed about whether AI is reliably superior at ideation, whether the institution of science should shift to more AI-led work, and whether humanity should strive for fully autonomous AI scientists. These disagreements echo throughout the scientific landscape, with recent back-to-back Nature pieces on both the capabilities[@yamada2026automation] and limitations[@nature2026humans] of science with fewer humans and more AI. These limitations became the focus of Week 3 of the Fall 2026 Collective [Artificial] Intelligence Seminar.