```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("week-summary/R/week-summary.R")
source("literature-review/R/literature-review.R")
```

# Literature Review and Synthesis: Using AI {#sec-literature-review-using}

::: {.callout-note icon="false" title="This content was identified, synthesized, and written by humans."}
📚 But we also like this AI-generated review (TBD) of the benefits and drawbacks of AI for literature review and synthesis.
:::

LLMs have an unprecedented capacity to conduct and assist with literature reviews — making them arguably one of the most essential additions to the modern-day research workflow (@fig-week-04-summary). This week, an in-class demo focused on how the "days of research assistants, Google Alerts, and journal subscriptions" are being replaced by AI-assisted workflows (@fig-week-04-literature-review). For example, armed with markdown files for organizing search strategies, key insights, and overarching project goals, one demo deployed Claude[@anthropic2026claude] to (a) search, screen, and organize resources, (b) extract key insights, connections, and themes, (c) brief the PI on recent developments, and (d) adjust the project vision based on PI feedback (@fig-week-04-literature-review, Panel B). This process runs automatically each week, allowing this researcher to sort through an increasingly growing and unmanageable scientific literature with more efficiency than ever before.

```{r}
#| label: fig-week-04-summary
#| fig-cap: "Overview of Week 4 of the Fall 2026 Collective [Artificial] Intelligence Seminar. Left: top discussion points, along with estimated level of agreement among seminar participants. Right: three most popular resources identified by seminar participants."
#| fig-width: 12
#| fig-height: 5.5
#| fig-align: center
week_summary_plot(week = 4, fig_width = 12, fig_height = 5.5)
```

From custom workflows (@fig-week-04-literature-review) to LLM skills[@wu2026skills] to off-the-shelf tools like Consensus[@consensus2026], resources compiled by seminar participants painted an overwhelming picture: AI *already* has revolutionized the literature review space. For example, an early evaluation demonstrated that LLMs are capable of reviewing up to 110 research records per minute, often achieving accuracy comparable to human coders who need months for the same task[@matsui2024screening]. However, not all models are created equal — some performing at human-level accuracy and others landing at the floor[@li2024screening]. This has led to new programs of research dedicated to optimizing these AI systems. For example, AI agents are capable of performing more accurate literature reviews when allowed to work in teams[@hilkenmeier2026agreement] — an insight that parallels many of the collective intelligence themes reviewed in this seminar[@Coles2022; @Coles2026; @nrc2015team] (see [Preface](../index.qmd)).

## Areas of consensus — and dissensus

Working in a team setting, of course, requires navigating both consensus and dissensus[@Coles2024; @dalkey1963delphi]. In terms of consensus, there was strong agreement that AI tools are ubiquitous in literature review, with browser-based applications like Consensus[@consensus2026], Gemini Notebook[@google2026notebook], and ChatGPT[@openai2026chatgpt] providing non-technical entry points for nearly all researchers. Furthermore, there was agreement that such workflows are still relatively novel, indicating a strong need for continued evaluation and reflection. As part of this reflection, we reached our most poignant disagreements. Are systematic reviews still noteworthy contributions to science? Are we ready to rely upon these tools in our own research? And, if so, what's left for humans?

"What's left for humans?" was not an abstract question. Nearly every seminar participant considered literature reviews an essential part of their initial training. Yet, many also acknowledged that they've largely stopped offering these same training opportunities in their own labs. In the ultra-competitive and fast-moving world of science[@bornmann2021growth], many researchers simply feel they cannot afford to lose the speed advantage of LLM-assisted literature review. Seminar participants had mixed emotions[@Larsen2017] about these developments. For example, some seminar participants noted the extremely unpleasant nature of a literature review assignment. One participant likened the notion of having a human spend months on a literature review task an LLM can complete in hours to "academic hazing". Some suggested it's a necessary skill to develop, but not to personally practice – akin to learning long division before being handed a calculator. Yet, even if that is the case, a philosophical tension remained. Every academic in the seminar has been collectively entrusted by society to understand, maintain, and develop the scientific record. What does it mean when we are increasingly disengaging from that literature directly? To what extent had we already done so far before the arrival of AI? Themes such as this became the focus of the following week of the Fall 2026 Collective \[Artificial\] Intelligence Seminar (see @sec-literature-review-critiquing).

```{r}
#| label: fig-week-04-literature-review
#| fig-cap: "A seminar participant shared an AI-assisted workflow for literature review and synthesis (Panel A) and demonstrated it running against a working project (Panel B). The agent searches and screens against a written strategy, retrieves and organizes the papers, extracts insights, connections, and themes into separate files, and then talks the results through with the principal investigator - who updates the project's guiding documents before the loop repeats."
#| fig-width: 12
#| fig-height: 4.59
#| fig-align: center
literature_review_plot()
```