```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("figure-packages/week-summary/R/week-summary.R")
source("figure-packages/data-collection/R/data-collection.R")
```

# Data Collection and Generation: Using AI {#sec-data-collection-using}

::: {.callout-note icon="false" title="This content was identified, synthesized, and written by humans."}
🤖 Nonetheless, we acknowledge that LLMs may have been able to roughly simulate how each of us would have responded in class.
:::

For researchers, advancements in AI have created entirely new modes for collecting and generating data. This not only includes AI-powered tools for extracting information at scale[@hill2026thematic], but also the use of AI agents to manipulate[@ogrady2025reddit] and simulate human behavior[@ashokkumar2026predict] (@fig-week-06-summary). As a concrete example of such possibilities, one seminar participant reviewed a grant proposal they cynically described as "social science for AI" (@fig-week-06-data-collection, Panel A). More specifically, they argued that some interesting questions — like "how well can we predict changes in happiness around the world?" — can now be comprehensively investigated with little-to-no human involvement (@fig-week-06-data-collection, Panel B). At least theoretically, AI tools can (1) compile the literature and derive hypotheses, (2) identify hypothesis-relevant data sources, (3) download and interoperate those data sources, and (4) deploy automated machine learning to establish initial estimates of the upper-bound of predictability. The result, they joked, is that they its easier than ever to "publish fast, die unhappy anyways". Whether that general sentiment is warranted and/or inevitable served as the focus of Week 6 of the Fall 2026 Collective \[Artificial\] Intelligence Seminar.

```{r}
#| label: fig-week-06-summary
#| fig-cap: "Overview of Week 6 of the Fall 2026 Collective [Artificial] Intelligence Seminar. Left: top discussion points, along with estimated level of agreement among seminar participants. Right: three most popular resources identified by seminar participants."
#| fig-width: 12
#| fig-height: 5.5
#| fig-align: center
week_summary_plot(week = 6, fig_width = 12, fig_height = 5.5)
```

Unique to this week's discussion was the introduction of "silicon sampling": using LLMs as substitutes for human participants. This technique is based on early evidence that LLMs can simulate output with many statistical similarities to real human responses[@argyle2023out]. Perhaps most impressively, a recent report indicated that GPT-4 could successfully simulate the results of 70 survey-based experiments conducted on nearly 120,000 participants[@ashokkumar2026predict]. When one considers the immense amount of resources that went into those original studies, silicon sampling presents a massive potential cost savings. Such results are suggestive of a future where researchers use silicon samples to (1) pilot ideas before committing to costly experiments, (2) more rapidly test ideas, and (3) replace human experimentation altogether (e.g., in contexts where we cannot ethically or feasibly run experiments with humans).

## Areas of consensus — and dissensus

Like past weeks, there was an intentional attempt to identify both areas of agreement and disagreement[@Coles2024; @dalkey1963delphi]. Most seminar participants agreed that at least basic data extraction tasks could be completely handed off to AI tools. We further agreed that there are poignant methodological challenges in instances where constructs-of-interest are subjective. For example, if humans don't agree on which verbal displays are appropriate in a given context[@matsumoto1990], there is not one but *multiple* ground truths to validate against. Furthermore, evidence that silicon samples can accurately replicate human responses in one context doesn't mean that they do so in the next[@yarkoni2022].

We most notably disagreed in conversations about how far things could and would go. For example, some seminar participants viewed silicon samples as a regular piloting technique in future research — others more so likened them to the popularity of existing simulation tools (e.g., computational modeling). We reflected upon the value of hard-earned scientific insights in a scenario where those patterns were predictable all along[@corneille2023]. And we stared in the face of a real-world dual use case: a public GitHub repository in which the same LLMs that may help us one day understand global happiness were subjected to an "AI torture chamber"[@cuthbertson2026torture]. Issues such as these set the backdrop for the following week of the Fall 2026 Collective \[Artificial\] Intelligence Seminar (see @sec-data-collection-critiquing).

```{r}
#| label: fig-week-06-data-collection
#| fig-cap: "A seminar participant shared a grant proposal they cynically described as \"science with AI in five easy steps.\" Panel A: the proposed workflow, from compiling the literature to a researcher who publishes fast and dies unhappy anyways. Panel B: the title and abstract of the grant, which proposes to predict changes in societal happiness with an AI multi-agent system."
#| fig-width: 12
#| fig-height: 5.33
#| fig-align: center
data_collection_plot()
```

##